---
title: "Growth Curve Report"
author: "Randy L Coryell"
date: "June 22, 2016"
output: pdf_document
toc: yes
toc_depth: 4
---


## Overview of Growth Study

This report will highlight data aquired during a 8h growth study to "investigate the neutrality of *Tn7* antibiotic (CAM) neutral marker insertion in *Vibrio fischeri* 

>###Strains:

>>>>ES114

>>>>ET401

>>>>EM17

>>*Note* All cultures were derived from a single colony, isolated on CAM plates (2.5ug/mL) and grown in liquid SWT (100mL) overnight. These overnight cultures were used to innoculate fresh SWT (100mL) media and grown at 225rpm and 28C for 1.5h. After taking a turbity measurement via spectrophotometry, each strain was diluted by placing a volume of this 1.5h culture into 100ml of fresh SWT. An initial OD was taken and the cultures were then placed into a shaking incubator @ 28C and 225RPM for the study.


***
> I want to walk down the street and have people say 'There goes Roy Hobbs, the best to ever play the game'.

>> Robert Redford, The Natural

***


## Installation of R packages 
Certain R packages, available on the CRAN repository, are required for this analysis.

```{undefinedecho=FALSE}
install.packages(c('reshape2', 'dplyr', 'ggplot2'))

```


## Raw Data
In order to generate growth curves from the raw data, it first must be recorded as a CSV file in MSEXCEL and then loaded into memory in R.

``` {r  read the raw CSV file}
rawdata <- read.csv("/Users/randycoryell/Growth Curve/Growth Curve/06212016_CAM_VS_NOCAMcurve.csv")

print(rawdata)

```


## Tidy the data

A reshaping of the data matrix is required in order to futher analyze the data set.


```{r tidy the data}
library(reshape2)

reshaped <- melt(rawdata, id=c("Time", "Temp"), variable.name="Strain",
value.name="OD600")
```
 
 Summary statistics can be called upon for the reshaped data

``` {r  Summary Stats for reshaped data}
summary(reshaped)
```

Now we need to take a quick look at the reshaped matrix to check for farmatting errors

```{r  reshaped error check}
head(reshaped)
```

When the data is replicated, there would need to be a nesting or grouping performed, this step would be useful for replicates in experimental evolution studies

```{r group by strain, eval=FALSE, include=TRUE}
grouped <- group_by(reshaped, Strain)

```

Sice these are single reads from the spectrophotometer, no confidence interval can be established, however an approximate growth curve can be produced.
``` {r  growth curve from single observations, fig=TRUE}
library(ggplot2)
print(ggplot(data=reshaped, aes(x=Time/60, y=OD600, color=Strain)) + geom_line() + labs(x="Time (Hours)", y="Absorbance at 600 nm"))

```